---
title: "Lab 1"
author: "Hanin Almodaweb"
format: html
theme: journal
embded-resources: true
fig-width: 6
fig-height: 4
---

## Question 1
1. Based on the help file, how many rows and how many columns does the `datasaurus_dozen` file have? What are the variables included in the data frame? Add your responses to your lab report, with relevant code in the associated R code chunk, and free-form text outside of the code chunk.

```{r}
library(datasauRus)
nrow(datasaurus_dozen)
ncol(datasaurus_dozen)
str(datasaurus_dozen)
```

There are 1846 rows and 3 columns in the datasaurus_dozen file. The variables included in the data frame are `dataset,` `x,` and `y.`

## Question 2
2. Plot `y` vs. `x` for the `dino` dataset. Then, calculate the correlation coefficient between `x` and `y` for just this dataset.

```{r}
dino_data <- datasaurus_dozen[datasaurus_dozen$dataset == 'dino', ]
plot(dino_data$x, dino_data$y)
cor(dino_data$x, dino_data$y)
```


## Question 3
3. Plot `y` vs. `x` for the `star` dataset, another one of the `datasaurus_dozen`. You can (and should) re-use code we introduced above, just replace the dataset name with the desired dataset. Then, calculate the correlation coefficient between `x` and `y` for this dataset. How does this value compare to the `r` of `dino`?

```{r}
star_data <- datasaurus_dozen[datasaurus_dozen$dataset == 'star', ]
plot(star_data$x, star_data$y)
cor(star_data$x, star_data$y)
```

The correlation coefficient for the `star` dataset (-0.0629611) is slightly less negative than that for the `dino` dataset (-0.06447185), indicating a marginally weaker negative linear relationship between x and y in the `star` dataset, though both correlations are very close to zero and suggest minimal linear association.

## Question 4
4. Plot `y` vs. `x` for the `circle` dataset. You can (and should) reuse code we introduced above, just replace the dataset name with the desired dataset. Then, calculate the correlation coefficient between `x` and `y` for this dataset. How does this value compare to the `r` of `dino`?

```{r}
circle_data <- datasaurus_dozen[datasaurus_dozen$dataset == 'circle', ]
plot(circle_data$x, circle_data$y)
cor(circle_data$x, circle_data$y)
```

The correlation coefficient for the `circle` dataset (-0.06834336) is slightly more negative than that for the `dino` dataset (-0.06447185), indicating a marginally stronger negative linear relationship between `x` and `y` in the `circle` dataset.

## Question 5
5. Plot all 13 datasets at once. 

```{r, fig.width=12, fig.height=10}
layout(matrix(1:16, nrow=4, ncol=4))
for(name in unique(datasaurus_dozen$dataset)){
  subset <- datasaurus_dozen[datasaurus_dozen$dataset == name, ]
  plot(subset$x, subset$y, main = name)
}
layout(1)
```

## Question 6
6. Calculate the correlation between the `x` and `y` variables for all 13 datasets. 

```{r}
sapply(unique(datasaurus_dozen$dataset), function(name){
    subset <- datasaurus_dozen[datasaurus_dozen$dataset == name, ]
    return(cor(subset$x, subset$y))
})
```

---
